---
title: "Integer Programming Assignment"
author: "Priyanka Jonnala"
date: "2023-11-20"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
Installing and loading the required libraries
```{r}
#install.packages("lpSolve")
library(lpSolve)
#install.packages("lpSolveAPI")
library(lpSolveAPI)
```
This table below provides an estimate of the number of workers needed each day of the week.
```{r}
Day_of_the_week = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
Workers_required = c(18,27,22,26,25,21,19)
Worker_per_day = data.frame(Day_of_the_week, Workers_required)
Worker_per_day
```
Package handlers at AP are guaranteed a five-day work week with two consecutive days off. The base wage for the handlers is $750 per week. Workers working on Saturday or Sunday receive an additional $25 per day. The possible shifts and salaries for package handlers are:
```{r}
Shifts = c(1,2,3,4,5,6,7)
Days_off = c("Sunday & Monday", "Monday & Tuesday", "Tuesday & Wednesday", "Wednesday & Thursday", "Thursday & Friday", "Friday & Saturday", "Saturday & Sunday")
Wages = c(775, 800, 800, 800, 800, 775, 750)
shift_wage = data.frame(Shifts, Days_off, Wages)
shift_wage
```
Defining the decision variables:

w1 = no.of workers in shift1

w2 = no.of workers in shift2

w3 = no.of workers in shift3

w4 = no.of workers in shift4

w5 = no.of workers in shift5

w6 = no.of workers in shift6

w7 = no.of workers in shift7

The Objective function is

Z = 775(w1) + 800(w2) + 800(w3) + 800(w4) + 800(w5) + 775(w6) + 750(w7)

Constraints:

Sunday    : 0(w1)+1(w2)+1(w3)+1(w4)+1(w5)+1(w6)+0(w7) >= 18

Monday    : 0(w1)+0(w2)+1(w3)+1(w4)+1(w5)+1(w6)+1(w7) >= 27

Tuesday   : 1(w1)+0(w2)+0(w3)+1(w4)+1(w5)+1(w6)+1(w7) >= 22

Wednesday : 1(w1)+1(w2)+0(w3)+0(w4)+1(w5)+1(w6)+1(w7) >= 26

Thursday  : 1(w1)+1(w2)+1(w3)+0(w4)+0(w5)+1(w6)+1(w7) >= 25

Friday    : 1(w1)+1(w2)+1(w3)+1(w4)+0(w5)+0(w6)+1(w7) >= 21

Saturday  : 1(w1)+1(w2)+1(w3)+1(w4)+1(w5)+0(w6)+0(w7) >= 19

Non-negativity Constraints: w1,w2,w3,w4,w5,w6,w7 >= 0
```{r}
solvelp = make.lp(7, 7)
#Setting the objective function
set.objfn(solvelp, c(775, 800, 800, 800, 800, 775, 750))
#Setting the objective function to "min"
lp.control(solvelp, sense = 'min')
```
Adding constraints
```{r}
set.row(solvelp,1,c(0,1,1,1,1,1,0))
set.row(solvelp,2,c(0,0,1,1,1,1,1)) 
set.row(solvelp,3,c(1,0,0,1,1,1,1))
set.row(solvelp,4,c(1,1,0,0,1,1,1)) 
set.row(solvelp,5,c(1,1,1,0,0,1,1)) 
set.row(solvelp,6,c(1,1,1,1,0,0,1)) 
set.row(solvelp,7,c(1,1,1,1,1,0,0))
```
Adding the right hand side coefficients.
```{r}
rhs<-c(18,27,22,26,25,21,19)
set.rhs(solvelp,rhs)
```
Setting the Constraint types
```{r}
set.constr.type(solvelp,c(">=", ">=", ">=", ">=", ">=", ">=", ">="))
set.bounds(solvelp,lower = rep(0,7))
```
Giving row names
```{r}
lp.rownames <- c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")
dimnames(solvelp) = list(lp.rownames,(1:7))
```
Column names
```{r}
lp.colnames <- c("w1","w2","w3","w4","w5","w6","w7")
set.type(solvelp, columns = 1:7, type = "integer")
solvelp
```
Solving the LP Problem
```{r}
solve(solvelp)
```
Finding the objective function of the LP Problem
```{r}
get.objective(solvelp)
```

```{r}
round(get.variables(solvelp))
```
```{r}
get.constraints(solvelp)
```
```{r}
Availability = matrix(c(0,4,5,0,8,1,0,
                 0,0,5,0,8,1,13,
                 2,0,0,0,8,1,13,
                 2,4,0,0,8,1,13,
                 2,4,5,0,0,1,13,
                 2,4,5,0,0,0,13,
                 2,4,5,0,8,0,0),ncol=7,byrow=TRUE)

colnames(Availability) = c("Shift1", "Shift2", "Shift3", "Shift4", "Shift5", "Shift6", "Shift7")

rownames(Availability) = c('Sunday', 'Monday','Tuesday','Wednesday','Thursday','Friday','Saturday')

Availability
```
This shows the availability of workers each day.
```{r}
rowSums(Availability)
```
CONCLUSION

From the above  formulation of the problem we can observe that

The total cost is $25,675

The optimal no.of workers available each day is:

Sunday    : 18

Monday    : 27

Tuesday   : 24

Wednesday : 28

Thursday  : 25

Friday    : 24

Saturday  : 19

